---
title: "HW9"
author: "Seana Perkins"
date: "2023-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Part one: Data tidying

The avenger data consists of ten columns for each avenger called Death[1-5] to Return[1-5]

Get the data into a format where the five columns for Death[1-5] are replaced by two columnes: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the result `deaths`. 

Similarly, deal with the returns of characters.

```{r}

library(tidyverse)
av = read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
names(av)
av0 = av %>% filter(Name.Alias != "")
av1 = av0 %>% gather(key = time, value = death, c(11, 13, 15, 17, 19)) %>% 
  select(Name.Alias, time, death) %>%
  mutate(time = parse_number(time))
av2 = av0 %>% gather(key = time, value = return, c(12, 14, 16, 18, 20)) %>% 
  select(Name.Alias, time, return) %>%
  mutate(time = parse_number(time))
av.neat = left_join(av1, av2, by = c("Name.Alias", "time")) 
av0 = av0 %>% filter(Name.Alias != "Vance Astrovik")
av1 = av0 %>% gather(key = time, value = death, c(11, 13, 15, 17, 19)) %>% 
  select(Name.Alias, time, death) %>%
  mutate(time = parse_number(time))
av2 = av0 %>% gather(key = time, value = return, c(12, 14, 16, 18, 20)) %>% 
  select(Name.Alias, time, return) %>%
  mutate(time = parse_number(time))
av.neat = left_join(av1, av2, by = c("Name.Alias", "time"))
```

``` {r}

av.neat %>% group_by(time) %>% summarise(total.num = n(),
                                         real.death = sum(death == "YES"), 
                                         fake.death = sum(death == "NO"), 
                                         missing = sum(death == ""))
av.neat %>% filter(time == 1, death == "YES") %>% 
  summarise(total.num = n(),
            real.return = sum(return == "YES"), 
            fake.return = sum(return == "NO"), 
            missing = sum(return == ""))
```

## Part two: Validation

Each team member picks one of the statements in the FiveThirtyEight analysis and fact checks it based on the data. Use dplyr functionality whenever possible.

Upload your answers and the code to the repository. Discuss and refine answers as a team.

Seana - "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team." *This quote appears to be incorrect, as the below summary statistics say that there were 82 deaths. 

```{r}

av.neat$death <- as.factor(av.neat$death)
summary(av.neat$death)

```


Shelby - "Jocasta — an android based on Janet van Dyne and built by Ultron — who has been destroyed five times and then recovered five times." 

  This is true, Josasta is the only character to have died 5 times and also returned. 



```{r}
av.neat %>% filter(time=="5", death=="YES")

```

Greg - "Of the nine avengers we see on screen - Iron Man, The Hulk, Captain America, Thor, Hawkeye, Black Widow,
Scarlet Witch, Quicksilver, and The Vision - every single one of them has died at least once in thcourse of their time
Avenging in the comics."

  This is likely true, but the dataset doesn't seem to contain The Vision. In the movies we see him die, so given that
    every other Avenger mentioned is shown to have tied by the database, We can say with a degree of certainty that
    this is true.

```{r}
av.neat %>% filter(Name.Alias %in% c("Anthony Edward \"Tony\" Stark", "Robert Bruce Banner", "Steven Rogers",
                                     "Thor Odinson", "Clinton Francis Barton", "Natalia Alianovna Romanova",
                                     "Wanda Maximoff", "Pietro Maximoff", "The Vision"), time == 1)
```



## Submission for HW9

In Canvas, submit an RMarkdown file (and the knitted html document) with:

1. the names of all team members that worked on the document.
2. a link to the github repo you were working in.
3. code and answers to all of the questions.
4. you can upload the data file, but you don’t have to